---
title: earthtones
output:
  md_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit this file directly. -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/",
  message = FALSE,
  warning = FALSE
)
```

[![CRAN\_Status\_Badge](http://www.r-pkg.org/badges/version/earthtones)](https://cran.r-project.org/package=earthtones) 
[![R-CMD-check](https://github.com/traitecoevo/earthtones/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/traitecoevo/earthtones/actions/workflows/R-CMD-check.yaml)

## Overview

**earthtones** is an R package designed to extract color palettes from remote imagery, providing visually appealing earth-inspired color schemes.

## Table of Contents

- [Overview](#overview)
- [Installation](#installation)
- [Usage](#usage)
  - [Example: Grand Canyon](#example-grand-canyon)
  - [Example: The Bahamas](#example-the-bahamas)
  - [Example: Uluru, Australia](#example-uluru-australia)
- [Using Palettes in Your Own Visualizations](#using-palettes-in-your-own-visualizations)
  - [ggplot2 Example](#example-with-ggplot2)
- [Contributing](#contributing)
- [Citation](#citation)
- [License](#license)

## Installation

Install the released version from CRAN:

```{r}
# install.packages("earthtones")
```

Or the latest development version from GitHub:

```{r, message=FALSE}
# remotes::install_github("traitecoevo/earthtones")
```

Load the package:

```{r}
library("earthtones")
```

## Usage

`earthtones` provides the function `get_earthtones`, which:

1. Downloads remote imagery from a supported provider--now using esri as default.
2. Converts image colors into a perceptually uniform color space ([CIE LAB](https://en.wikipedia.org/wiki/Lab_color_space)).
3. Identifies dominant colors using clustering methods (`kmeans` or `pam`).
4. Returns an aesthetically pleasing color palette.

### Example: Grand Canyon

```{r grandcanyon}
get_earthtones(latitude = 36.094994, longitude = -111.837962, 
               zoom = 12, number_of_colors = 6)
```

Adjust `number_of_colors` to control how many dominant colors are returned. The `zoom` parameter controls the level of detail, with larger numbers providing closer views.

### Example: The Bahamas

Explore vibrant tropical colors, using the default kmeans clustering:

```{r bahamaskmeans}
get_earthtones(latitude = 24.2, longitude = -77.88, zoom = 11, number_of_colors = 5,
               method="kmeans")
```

or use a different clustering algorithm (Partitioning Around Medoids) to choose the colors from image:

```{r bahamaspam}
get_earthtones(latitude = 24.2, longitude = -77.88, 
               zoom = 11, number_of_colors = 5,
              method="pam")
```

### Example: Uluru, Australia

Earth tones from the iconic Australian landscape:

```{r uluru}
get_earthtones(latitude = -25.33, longitude = 131.0396, zoom = 10, number_of_colors = 4)
```

## Using Palettes in Your Own Visualizations

To retrieve the color palette without plotting the map, set `include.map = FALSE`:

```{r}
palette <- get_earthtones(latitude = 24.2, longitude = -77.88,
                          zoom = 11, number_of_colors = 5,
                          include.map = FALSE)

print(palette)
```

This returns a standard R color palette ready to be used directly in your visualizations or designs.

### Example with ggplot2

Here's how you might apply the palette to a `ggplot2` visualization:

```{r ggplot_example}
library(ggplot2)

# Generate the palette
bahamas_palette <- get_earthtones(latitude = 24.2, longitude = -77.88,
                                  zoom = 11, number_of_colors = 3,
                                  include.map = FALSE)

# Example ggplot
ggplot(mtcars, aes(factor(cyl), fill = factor(cyl))) +
  geom_bar() +
  scale_fill_manual(values = bahamas_palette) +
  theme_minimal() +
  labs(title = "MTCARS Cylinders with Bahamas Earthtones",
       fill = "Cylinders")
```

## Contributing

Contributions are welcome! Please report issues or suggest features on the [GitHub issues page](https://github.com/traitecoevo/earthtones/issues). Pull requests are also encouraged.

## Citation

If you use **earthtones** in your research or projects, please cite it using:

```{r}
citation("earthtones")
```

## License

This package is released under the MIT License.
